A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2015-10-24 imaging_demo/ms.tex
Add content under section 6
From 68b60bc (master):
=====================

In ms.tex,

(*) Add content under section 6

In the most general case, this gridding kernel could contain information on the
$w$-projection effect, and even other time-dependent ionospheric effects. For a
stationary antenna array in the absence of any time-dependent effects, this
mapping must only be determined once in the antenna array coordinate frame. The
antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described
as a transformation matrix from the space of measured electric fields by the antennas
($\mathbf{a}$) to the antenna array grid denoted by the coordinate $\mathbf{r}$. Since each
antenna occupies a footprint typically the size of its aperture,
$\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
$N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
block. Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
matrix and the grid containing the mapped aperture footprints of the antennas.

@@

The antenna array aperture illumination over the entire grid,
$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the individual
antenna aperture illuminations:
\begin{align}\label{eqn:gridding-convolution}
  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
\end{align}
where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved using
efficient multiplication with the sparse matrix created in the antenna-to-grid
mapping process. Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
effects of the ionosphere or the instrument, it needs to be computed just once
for the entire observation. However, the gridding of electric fields must be
computed at every readout of the electric field spectra,
\begin{align}
  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
\end{align}
% The sentence below was already stated earlier.
%In practice the grid is set to be a power of 2 such that the grid spacing is at
%most $\lambda/2$ even at the highest frequency in the observing band.

@@

We study the effect of the differences in gridded weights on the image plane. 
The synthesized beam can be written as:
\begin{align}\label{eqn:syn-beam}
  B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
  \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
\end{align}
where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
$\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
If the errors in the determination of $r_x$ and $r_y$ are denoted by 
$\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
estimated using a simple propagation of errors, where,
\begin{align}
  \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
  &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
  &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
  \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}). \label{eqn:syn-beam-pdev2}
\end{align}
It must be noted that although we start with a squaring operation in 
equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
\ref{eqn:syn-beam-pdev2}, the summations are performed only when 
$\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
\begin{align}
  \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
  &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
  &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
\end{align}
For the current simulations, 
\begin{align}
  \delta r_x = \delta r_y &\simeq 1, \\
  \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
  \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
\end{align}
Hence,
\begin{align}
  \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
\textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
\end{align}
where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
the error on the synthesized beam is proportional to the antenna power pattern
and increases linearly outward in radial bins of direction cosines.

Fig.~\ref{fig:psf-diff} shows the difference between the synthesized beams 
obtained with the two methods. A difference map between the two synthesized 
beams is shown on top. The amplitude of the difference appears to be 
modulated by the directional voltage response of the antenna. At the bottom, 
in radial bins, the rms of the synthesized beam (gray) and the rms of the 
difference map (black) are plotted in percentage units relative to the peak 
(to be read using the axis on the left side of the plot). The antenna power 
pattern (red; to be read using the scale on the right) is plotted for reference.

@@

The synthesized beam rms is proportional to the antenna power pattern as 
expected from a point spread function of uncorrected for the antenna power 
pattern. A notable observation is that the rms of the difference of
the synthesized beams seems proportional to the directional antenna voltage
pattern and also increases gradually with distance from the center as seen 
from the diverging amplitudes between the black and red curves as predicted
in equation~\ref{eqn:syn-beam-error}. The rms of the difference appears to 
be bounded by the actual rms of the synthesized beam indicating that these
rounding errors in gridding do not affect the statistics of the images or
the synthesized beams.

@@

\begin{table}
  \centering
  \caption{Radio telescopes and array layouts.}
  \label{tab:antenna-layouts}
  \begin{threeparttable}
  \begin{tabular}{ccccc} 
    \hline
    Telescope & Core & Number & Antenna & Frequency \\
              & size & of Antennas & size & \\
    (key)     & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
    \hline
    MWA & 1400 & 112 & 16 & 150 \\
    (MWA-112) \\ \\
    MWA & 1400 & 496 & 16 & 150 \\
    (MWA-496) \\ \\
    LOFAR & 3500 & 24 & & 50 \\
    (LOFAR-C) & & & & \\ \\
    LWA1 & 100 & 256 & 10 & 50 \\
    (LWA1) \\ \\
    LWA-OV\tnote{a} & 200 & 256 & 10 & 50 \\
    (LWA-OV) & & & & \\ \\
    HERA & 70 & 19 & 154 & 150 \\
    (HERA-19) \\ \\
    HERA & 98 & 37 & 154 & 150 \\
    (HERA-37) \\ \\
    HERA & 294 & 331 & 154 & 150 \\
    (HERA-331) \\ \\
    HERA & 1330 & 6769\tnote{b} & 154 & 150 \\
    (HERA-6769) \\ \\
    SKA1-Low & 1000 & 750\tnote{c} & 962 & 150 \\
    (SKA1-LC) \\ \\
    SKA1-Low & 1000 & 192,000\tnote{c} & 962 & 150 \\
    Dipoles\tnote{d} \\
    (SKA1-LCD) \\
    \hline
  \end{tabular}
  \begin{tablenotes}
    \item[a] Owens Valley LWA
    \item[b] Hypothetically chosen to have a total collecting area of 
      1~km$^2$
    \item[c] This is the number of elements expected to be in the core, 
      three-fourths of the total number
    \item[d] All dipoles are used as independent elements without station 
      beamforming
  \end{tablenotes}
  \end{threeparttable}
\end{table}
\footnotetext{LWA at Ovens Valley}

@@

Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the ratio of the
number of computations required with visibility based imaging relative to 
MOFF imaging and the boundaries where this ratio is unity. The color scale 
refers to the ratio while using a 1~m antenna element. Pairs of black and
white lines of the same line style correspond to a specific antenna size. For 
a given antenna size (annotated on the plot), the white line denotes the 
minimum baseline length that could achieve the closest packing as a function 
of the number of antennas. Regions to the right of the white line for a 
given antenna size imply a physically impossible scenario where antennas will 
have to be packed overlapping with each other. For the same antenna size, the 
black line denotes the boundary where the ratio of the number of computations 
with either algorithm becomes unity. Regions to the right of the black lines 
favour usage of the MOFF algorithm. Thus the wedge between the white and black 
lines denotes the region in parameter space where MOFF algorithm holds a 
definite advantage while visibility based algorithms will be favoured to the 
left of the black lines. As antenna size increases the maximum number of 
antennas for a dense packing as a function of baseline length decreases. Hence 
the white lines shift leftward as antenna size increases. Similarly, with 
increase in antenna size, $N_\textrm{g}$ also decreases when field of view 
imaging is achieved with an increasing grid spacing equal to antenna size and
hence lowers the amount of computations required with the MOFF algorithm. This
shifts the black curves leftward. HERA-331, HERA-1027, LWA1, SKA1-LC and 
SKA1-LCD are inside their respective wedges indicating they are in regions that
favour the usage of MOFF imaging.

@@

\section{Conclusions}\label{sec:conclusions}
\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

@@

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

>>>>

In the most general case, this gridding kernel could contain information on the
$w$-projection effect, and even other time-dependent ionospheric effects. For a
stationary antenna array in the absence of any time-dependent effects, this
mapping must only be determined once in the antenna array coordinate frame. The
antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described
as a transformation matrix from the space of measured electric fields by the 
antennas ($\mathbf{a}$) to the antenna array grid denoted by the coordinate 
$\mathbf{r}$. Since each antenna occupies a footprint typically the size of its 
aperture, $\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
$N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
block. This sparse matrix is stored in a Compressed Sparse Row (CSR) format. 
Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
matrix and the grid containing the mapped aperture footprints of the antennas.

@@

The antenna array aperture illumination over the entire grid,
$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the individual
antenna aperture illuminations:
\begin{align}\label{eqn:gridding-convolution}
  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
\end{align}
where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved by
efficient multiplication with the sparse matrix created in the antenna-to-grid
mapping process using the sparse matrices module in Python's SciPy package. 
Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
effects of the ionosphere or the instrument, it needs to be computed just once
for the entire observation. However, the gridding of electric fields must be
computed at every readout of the electric field spectra,
\begin{align}
  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
\end{align}
% The sentence below was already stated earlier.
%In practice the grid is set to be a power of 2 such that the grid spacing is at
%most $\lambda/2$ even at the highest frequency in the observing band.

@@

% The synthesized beam can be written as:
% \begin{align}\label{eqn:syn-beam}
%   B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
%   \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
% \end{align}
% where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
% $\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
% If the errors in the determination of $r_x$ and $r_y$ are denoted by 
% $\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
% estimated using a simple propagation of errors, where,
% \begin{align}
%   \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
%   &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
%   &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
%   \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}). \label{eqn:syn-beam-pdev2}
% \end{align}
% It must be noted that although we start with a squaring operation in 
% equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
% \ref{eqn:syn-beam-pdev2}, the summations are performed only when 
% $\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
% auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
% \begin{align}
%   \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
%   &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
%   &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
% \end{align}
% For the current simulations, 
% \begin{align}
%   \delta r_x = \delta r_y &\simeq 1, \\
%   \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
%   \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
% \end{align}
% Hence,
% \begin{align}
%   \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
% \textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
% \end{align}
% where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
% the error on the synthesized beam is proportional to the antenna power pattern
% and increases linearly outward in radial bins of direction cosines.

We study the effect of the differences in gridded weights on the image plane. 
Fig.~\ref{fig:psf-diff} shows the difference between the synthesized beams 
obtained with the two methods. A difference map between the two synthesized 
beams is shown on top. The amplitude of the difference appears to be 
modulated by the directional power response of the antenna. At the bottom, 
in radial bins, the rms of the synthesized beam (gray) and the rms of the 
difference map (black) are plotted in percentage units relative to the peak 
(to be read using the axis on the left side of the plot). The antenna power 
pattern (red; to be read using the scale on the right) is plotted for reference.

@@

The synthesized beam rms is proportional to the antenna power pattern as 
expected from a point spread function uncorrected for the antenna power 
pattern. The amplitude of the rms of difference of the synthesized beams is 
also modulated by the antenna power pattern. The rms of the difference is 
definitely lesser than the rms of the synthesized beam in the central regions 
up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This implies that the beams are well 
matched in the central regions. In the outer regions, their mismatch is 
comparable to the rms of synthesized beams. This indicates the two 
synthesized beams are not completely randomly different from each other in
which case the rms of the difference would have been $\approx \sqrt{2}$
higher than the rms of the each of the synthesized beams. This indicates 
that while differences exist, large fractions of them are still well matched
to each other.Thus the rounding errors in gridding do not affect the 
statistics of the images or the synthesized beams.

@@

\begin{table}
  \centering
  \caption{Radio telescopes and array layouts.}
  \label{tab:antenna-layouts}
  \begin{threeparttable}
  \begin{tabular}{ccccc} 
    \hline
    Telescope & Core & Number & Antenna & Frequency \\
              & size & of Antennas & size & \\
    (key)     & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
    \hline
    MWA & 1400 & 112 & 16 & 150 \\
    (MWA-112) \\ \\
    MWA & 1400 & 496 & 16 & 150 \\
    (MWA-496) \\ \\
    LOFAR & 3500 & 24 & & 50 \\
    (LOFAR-C) & & & & \\ \\
    LWA1 & 100 & 256 & 10 & 50 \\
    (LWA1) \\ \\
    LWA-OV\tnote{a} & 200 & 256 & 10 & 50 \\
    (LWA-OV) & & & & \\ \\
    HERA & 70 & 19 & 154 & 150 \\
    (HERA-19) \\ \\
    HERA & 98 & 37 & 154 & 150 \\
    (HERA-37) \\ \\
    HERA & 294 & 331 & 154 & 150 \\
    (HERA-331) \\ \\
    HERA & 1330 & 6769\tnote{b} & 154 & 150 \\
    (HERA-6769) \\ \\
    SKA1-Low & 1000 & 750\tnote{c} & 962 & 150 \\
    (SKA1-LC) \\ \\
    SKA1-Low & 1000 & 192,000\tnote{c} & 962 & 150 \\
    Dipoles\tnote{d} \\
    (SKA1-LCD) \\
    \hline
  \end{tabular}
  \begin{tablenotes}
    \item[a] Owens Valley LWA
    \item[b] Hypothetically chosen to have a total collecting area of 
      1~km$^2$
    \item[c] This is the number of elements expected to be in the core, 
      three-fourths of the total number
    \item[d] All dipoles are used as independent elements without station 
      beamforming
  \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the ratio of the
number of computations required with visibility based imaging relative to 
MOFF imaging and the boundaries where this ratio is unity. The color scale 
refers to the ratio while using a 1~m antenna element. Pairs of black and
white lines of the same line style correspond to a specific antenna size. For 
a given antenna size (annotated on the plot), the white line denotes the 
minimum baseline length that could achieve the closest packing as a function 
of the number of antennas. Regions to the right of the white line for a 
given antenna size imply a physically impossible scenario where antennas will 
have to be packed overlapping with each other. For the same antenna size, the 
black line denotes the boundary where the ratio of the number of computations 
with either algorithm becomes unity. Regions to the right of the black lines 
favour usage of the MOFF algorithm. Thus the wedge between the white and black 
lines denotes the region in parameter space where MOFF algorithm holds a 
definite advantage while visibility based algorithms will be favoured to the 
left of the black lines. As antenna size increases the maximum number of 
antennas for a dense packing as a function of baseline length decreases. Hence 
the white lines shift leftward as antenna size increases. Similarly, with 
increase in antenna size, $N_\textrm{g}$ also decreases when field of view 
imaging is achieved with an increasing grid spacing equal to antenna size and
hence lowers the amount of computations required with the MOFF algorithm. This
shifts the black curves leftward. HERA-331, HERA-1027, LWA1, SKA1-LC and 
SKA1-LCD are inside their respective wedges indicating they are in regions that
favour the usage of MOFF imaging. For fixed baseline length, regions favouring
the MOFF algorithm tend to be towards large $N_\textrm{a}$ indicating large-N
dense array layouts with smaller antenna elements are best suited for deploying 
the MOFF algorithm.

@@

\section{Conclusions}\label{sec:conclusions}

As radio astronomy is entering a new era, advances in instrumentation have to
be accompanied by equal advances in processing techniques to manage 
computational resources. Many future radio telescopes such as the SKA, HERA and 
LWA are headed towards the large-N dense array layout model for which 
computational cost from traditional FX/XF correlator based architecture and 
visibility based imaging starts rising steeply. We have provided the first 
software demonstration of a general purpose imaging algorithm using our generic 
and efficient EPIC software that is designed to bring this cost down from 
$\mathcal{O}(N^2)$ to $\mathcal{O}(N\log N)$. Under the class of direct imaging 
techniques, ours is one of the most generic -- neither does it place any 
constraint on the array layout to be on a regular grid nor does it require the
antenna array to be homogeneous.

Our package, now publicly available, written in object oriented Python is 
highly modularized and parallelizable. It includes an implementation of the 
MOFF algorithm in addition to visibility based software holography imaging and 
a data simulator for sky models. It has been successfully tested on simulated 
as well as real LWA observations.

The MOFF algorithm packaged with EPIC is already found to be most suitable 
for many present and planned radio telescopes such as the LWA, HERA and SKA. 
In general, MOFF is most suited to operate in the region of parameter space
characterized by dense packing of a large number of antennas especially when 
consisting of a large number of small antenna elements.

A unique and significant advantage is the instantaneous availability of 
calibrated time-domain images bundled together with the hardware frontend 
such as the F-engine as an integrated module. Hence, it is a compelling 
candidate for time-domain radio astronomy, e.g. search for and monitoring of 
transients. Transient detection pipeline at the backend of EPIC can be 
fine-tuned to target fast transients such as the Fast Radio Bursts 
\citep[FRB;][]{tho13} on millisecond timescales at GHz frequencies or slow 
transients from planetary and exoplanetary origins at frequencies around 
100~MHz.

Thus, EPIC with the MOFF algorithm packaged is uniquely poised to offer a
substantial advantage to imaging with large-N dense arrays typical of 
next-generation radio telescopes as well as push the frontiers of 
time-domain astronomy to fill gaps in understanding the science behind 
phenomena responsible for extreme transient events in the Universe.

In the near future, we plan to upgrade our current Python implementation 
of EPIC to a GPU based pipeline in order to operate on real-time data and 
develop a transient trigger and monitor backend. In the meanwhile, we plan 
to demonstrate the capability of EPIC to calibrate and image from heterogeneous 
arrays and incorporate corrections for non-coplanarity of baselines and 
direction-dependence of calibration.

\section*{Acknowledgements}

We thank Larry D'Addario, Gregg Hallinan and Harish Vedantham for their valuable 
inputs.

@@

% \appendix

% \section{Some extra material}

% If you want to present additional material which would interrupt the flow of the main paper,
% it can be placed in an Appendix which appears after the list of references.
